Reducing Acceptance Marks in Emerson-Lei Automata by QBF-solving

### 2022-02-26 paper/main.tex
progress in section 3

We start with several simple observations. Acceptance of a run $\pi$
is fully determined by the set $\minf(\pi)$.
% of acceptance marks that
% appear infinitely often on $\pi$, which is in turn fully determined by
% the set $\minf(\pi)$ of transitions appearing infinitely often in
% $\pi$.
Further, each set $\minf(\pi)$ for a run $\pi$ is a cycle and
vice versa. Hence, our reduction methods is looking for a new
acceptance formula $\psi$ and function $\rem$ such that for each cycle
$T\subseteq\delta$, it holds that $T$ satisfies $\varphi$ if and
only if $\rem(T))$ satisifes $\psi$. This can be formally written
by the formula
\[
  \forall T\subseteq\delta~.~\mcycle(T)\implies(T~\msat~\varphi ~\iff~ \rem(T)~\msat~\psi).
\]

@@

@@

This section explains the construction of QBF formulas for a given
TELA $\mA =(Q,M,\Sigma,q_I,\delta,\varphi)$ and two integer parameters
$C,K\ge 0$ such that satisfiability of such a formula implies that
there exists an acceptance formula in DNF with $C$ clauses and $K$
acceptance marks and a placement of these marks on automata
transitions such that $\mA$ with this acceptance formula and
acceptance marks and acceptance formula

>>>>

We start with several simple observations. Acceptance of a run $\pi$
is fully determined by the set $\minf(\pi)$.
% of acceptance marks that
% appear infinitely often on $\pi$, which is in turn fully determined by
% the set $\minf(\pi)$ of transitions appearing infinitely often in
% $\pi$.
Further, each set $\minf(\pi)$ for a run $\pi$ is a cycle and vice
versa. Hence, our reduction methods is looking for a new acceptance
formula $\psi$ and function $\rem$ such that for each cycle
$T\subseteq\delta$, it holds that $T$ satisfies $\varphi$ if and only
if $\rem(T)$ satisifes $\psi$. This can be roughly denoted by the
formula
\[
  \forall T\subseteq\delta~.~\mcycle(T)\implies\big(\msat_\varphi(T) ~\iff~ \msat_{\psi,\rem}(T)\big).
\]
In fact, this corresponds to the shape of the QBF formulas we will
construct. As we are looking for $\psi$ and $\rem$ such that the
formula holds, the subformula $\msat_{\psi,\rem}(T)$ contains many
free variables representing possible instances of $\psi$ and
$\rem$. If the formula is satisfiable, then its models encode desired
instances of $\psi$ and $\rem$.  In the following, we assume that we
are looking for a new acceptance formula $\psi$ in DNF. Note that our
reduction method can be easily adopted to look for $\psi$ of a
different shape.

Now we describe the construction of QBF formulas in detail. The
construction is parametrized by two integers $C,K\ge 0$, where $K$ is
the desired number of acceptance marks and $C$ is the number of
clauses of $\psi$. For simplicity, we assume that the reduced
automaton will use the acceptance marks $N_K=\{1,2,\ldots,K\}$. First
we desribe Boolean variables used in the constructed QBF formulas.
\begin{itemize}
\item The fact whether a transition $t$ is in the set $T$ is
  represented by a boolean variable $e_t$. By $\vec{e}$ we denote the vector
  of length $|\delta|$ of all variables $e_t$.
\item 
\end{itemize}

\[
  \forall \vec{e}~.~\mcycle(\vec{e})\implies\big(\msat_\varphi(\vec{e}) ~\iff~ \msat(\vec{e},\vec{k},\vec{p})\big).
\]

@@

\subsection{Lightweight version of $\mcycle(\vec{e})$ (level 1)}

\subsection{Balanced version of $\mcycle(\vec{e})$ (level 2)}

\subsection{Strict version of  $\mcycle(\vec{e})$ (level 3)}

@@

This section explains the construction of QBF formulas for a given3TELA $\mA =(Q,M,\Sigma,q_I,\delta,\varphi)$ and two integer parameters
$C,K\ge 0$ such that satisfiability of such a formula implies that
there exists an acceptance formula in DNF with $C$ clauses and $K$
acceptance marks and a placement of these marks on automata
transitions such that $\mA$ with this acceptance formula and
acceptance marks and acceptance formula
